---
title: "FTE Enrollment"
subtitle: "Full time equivalent enrollment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{full_time_equivalent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(utDataStoR)
library(readr)
```

Full-Time Equivalent (FTE) is a measure calculated on student credit hours and student level.

FTE for undergraduate students is based on 15 credit hours.

``` sql
`r display_sql_with_comments('fte', 'fte_ug_cred_hour_census.sql')`
```

FTE for graduate students is based on 10 credit hours.

``` sql
`r display_sql_with_comments('fte', 'fte_gr_cred_hour_census.sql')`
```

FTE is reported to USHE five times per fiscal year:

-   Summer End of Term
-   Fall 3rd Week Census
-   Fall End of Term
-   Spring 3rd Week Census
-   Spring End of Term

USHE summarizes FTE by credit hour type, residency, gender, and budget.

IPEDS calculates FTE in the 12 Month Enrollment survey based on a reporting calendar from July 1, 20XX - June 20YY. Utah Tech excludes Summer academic terms and reports Spring and Fall term census data.
