---
title: "FTE Enrollment"
subtitle: "Full time equivalent enrollment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{full_time_equivalent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utDataStoR)
library(readr)
```

Full-Time Equivalent (FTE) is a measure calculated on student credit hours and student level. 

FTE for undergraduate students is based on 15 credit hours. 

```sql
`r display_sql_with_comments('fte', 'fte_ug_cred_hour_census.sql')`
```

FTE for graduate students is based on 10 credit hours.

```sql
SELECT b.term_desc,
       b.season,
          SUM(a.attempted_credits) as attmp_cred,
          ROUND(SUM(a.attempted_credits)/10, 2) as attmp_cred_fte,
          a.course_level_id,
          a.version_desc
     FROM export.student_section_version a
LEFT JOIN export.term b
       ON a.term_id = b.term_id
    WHERE a.is_enrolled = TRUE
      AND a.version_desc = 'Census' --Version description of snapshot status
      AND a.course_level_id = 'GR'
      AND DATE_PART('year', NOW()) - b.academic_year_code :: INT <= 5
GROUP BY a.course_level_id, b.season, b.term_desc, a.version_desc;
```

FTE is reported to USHE five times per fiscal year:

- Summer End of Term
- Fall 3rd Week Census
- Fall End of Term
- Spring 3rd Week Census
- Spring End of Term

USHE summarizes FTE by credit hour type, residency, gender, and budget.

IPEDS calculates FTE in the 12 Month Enrollment survey based on a reporting calendar from July 1, 20XX - June 20YY. Utah Tech excludes Summer academic terms and reports Spring and Fall term census data.





