---
title: "Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(utDataStoR)
```

Groups define the ways students are categorized. These groups can be placed in several classifications.

# Demopgraphic Groups
Data in this classification is collected from the student at the time of their application for admission. This data is stored at the student level, in the student table. This data is doesn't change often, but can be pulled at Current, End of Term or Census.

## Gender
Gender is the biological sex on a student's birth certificate. The options are:

- Male 
- Female
- Unspecified

## IPEDS Race/Ethnicity
IPEDS Race/Ethnicity is data reporting the student's race and ethnicity. There are nine options:

- Hispanic
- American Indian/Alaskan
- Asian
- Black/African American
- Hawaiian/Pacific Islander
- White
- Multiracial
- Non-Resident Alien (for those with visa's)
- Unspecified

## Underrepresented Minority
Underrepresented Minority students are those students who are ...
HELP

## BIPOC
BIPOC students are those whose race/ethnicity is black, indigenous, or other people of color (Hispanic, Asian and multi-racial). The options are:

- BIPOC
- non-BIPOC

## Citizenship
Citizenship is the student's relationship to the United States. The options are: 

- US Citizen
- Non-Res Alien-Visa or temp bas
- Res Alien-immigrant status
- Non-Citizen National of US
- Non-immigrant Alien 53B-8-106
- Deferred Action
- Student curr defined as Other

## International Student
An international student is a student who is from a foreign country (has a citizenship code of Non_Res Alien-Visa) and comes to the US to take classes on a temporary F-1 visa. This is a subset of Non-Resident Alien students (see IPEDS Race/Ethnicity). If the sql output is TRUE then the student is an international student.

## Age Band
Age is how old a student is today based on their birth date. Reported as a grouping: 

- Less than 18
- 18-24
- 25-34
- 35-44
- 44-59
- 60 plus

## First Generation Student
A first generation student is a student where neither parent nor guardian has received their bachelor's degree. On the application the student is asked if either of their parent(s) or guardian(s) has received a bachelor degree, if they respond no then the student is a first generation student. If they respond yes or if they don't answer then the student is not a first generation student. If the sql output is TRUE then the student is a first generation student.

## Veteran
A veteran is a student who has served in the armed forces or is receiving veteran benefits (including family members). If the sql output is TRUE then the student is a veteran.

## Student Athlete
A student athlete is a student who was EVER on the official team roster on or after the first scheduled contest or date of NCAA competition. If the sql output is TRUE then the student is an athlete.

## First Admit Country
First admit country is the the country listed on the students first application where they were admitted.

## First Admit State
First admit state is the the state listed on the students first application where they were admitted. Students who were not in the US get grouped into Foreign State.

## First Admit County
First admit county is the the county listed on the student first application where they were admitted. Students who were not from Utah get grouped into Out of state, in US; or Out of state, out of US.

## ACT Composite Score
ACT Composite Score is the students highest ACT composite score.


## SQL for Demographic Groups

```sql
SELECT a.student_id,
       CASE
           WHEN a.gender_code = 'M' THEN 'Male'
           WHEN a.gender_code = 'F' THEN 'Female'
           ELSE 'Unspecified'
           END AS gender_desc,
       a.ipeds_race_ethnicity,
       a.is_underrepresented_minority,
       CASE
           WHEN a.is_american_indian_alaskan = 'TRUE' THEN 'BIPOC'
           WHEN a.is_asian = 'TRUE' THEN 'BIPOC'
           WHEN a.is_black = 'TRUE' THEN 'BIPOC'
           WHEN a.is_hawaiian_pacific_islander = 'TRUE' THEN 'BIPOC'
           WHEN a.is_hispanic_latino_ethnicity = 'TRUE' THEN 'BIPOC'
           WHEN a.is_multi_racial = 'TRUE' THEN 'BIPOC'
           ELSE 'non-BIPOC'
           END AS BIPOC_student,
       a.us_citizenship_code,
       a.us_citizenship_desc,
       a.is_international,
       CASE
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) < 18 THEN 'less than 18'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 18 and 24 THEN '18 to 24'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 25 and 34 THEN '25 to 34'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 35 and 44 THEN '35 to 44'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 45 and 59 THEN '45 to 59'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) > 59 THEN '60 plus'
              ELSE 'error'
          END AS age_band,
       a.is_first_generation,
       a.is_athlete,
       a.is_veteran,
       a.first_admit_country_desc,
       a.first_admit_state_desc,
       a.first_admit_county_desc,
       a.act_composite_score
  FROM export.student a
```


# Term Based Groups
Data in this classification can change each semester and is based on information from the system for this current term such as program, credit hours attempted, etc. This data is stored at the student term level, in the student_term_level table. This data changes often, and can be pulled at Current, End of Term or Census.

## College
College is an organizational structure that houses departments. Based where off of the primary program the student has declared, students will be assigned to departments that are assigned to colleges. There are eight options: 

- COB - College of Business
- COE - College of Education
- COA - College of the Arts
- GCO - College of Global and Community Outreach
- COHS - College of Health Sciences
- CHASS - College of Humanities and Social Sciences
- CSET - College of Science, Engineering and Technology
- UCOL - University College

## Department
Department is an organizational structure that houses faculty, courses, and programs. Based off of the primary program the student has declared, students will be assigned to departments based on their program.

## Program
Program is a area of study that includes the degree, major, and concentration (when applicable) choosen by a student. Programs are assigned to a department and departments are assigned to colleges.

## Concentration
Concentration is a specific area of emphasis or specialization in a major chosen by a student. Not all majors have a concentration option.

## Degree Desc
Degree description is the description of the degree id/code that is tied to a students program.

## Major Desc
Major description is the description of the major code that is tied to a students program.

## Degree Seeking Student
A student is a degree seeking student if the students primary program leads to a credential. If the sql output is TRUE then the student is a degree seeking student.

## Student Type
Student Type is a student's enrollment status or registration status for the term. The options are:

- H -	High School
- F	- Freshman
- C -	Continuing Registration
- R	- Readmit
- T -	Transfer
- P	- Personal Interest, Non-Degree
- 1	- New Graduate
- 2	- Transfer Graduate
- 3	- Readmit Graduate
- 5	- Continuing Graduate

## Student Level
A student's level is the level of the primary degree a student is pursing. The options are: Undergraduate or Graduate.

## Class Level
A student's class level is based on their overall cumulative earned credit hours. A student's class level is not the same as their year in school, it is based on credit hours. There can be sophomores in their first semester, as well as freshmen in their third year.

All graduate students are reported as Graduate.
For undergraduate students, the options are: 

- Freshman <30 credit hours
- Sophomore 30-59 credit hours 
- Junior 60-89 credit hours
- Senior 90 or more credit hours 

## Credit Load
Credit load is the status of a student's time commitment based off of how many credit hours attempted and the student's level for a given semester. The options are: 

- Full-time for Undergraduate students 12 or more credit hours, for Graduate students 9 or more credit hours
- Part-time for Undergraduate students less than 12 credit hours, for Graduate students less than 9 credit hours.

## Tuition Residency
Tuition Residency is the official status of student's residence for tuition purposes. The options are:

- 0 - Undeclared
- A - American Indian Tribal Resident
- B - Housebill 102 - 2023
- C - Changed from Non-Resident
- G - Good Neighbor Tuition
- H - Housebill 118/144 Waivers
- M - Military, Out-of-State
- N - Non-Resident of UT
- R - UT Resident
- S - Surcharge for excessive credits

## State Residency
State Residency summarizes tuition residency into two categories, in-state or out-of-state. The options are:

- In-state Residency
- Out-of-state Residency

## Global Residency
Global Residency first looks at international status, then looks at state residency. The options are:

- International
- In-State Residency
- Out-of-State Residency

## High School Student
A High School Student is a student is who is taking a course at Utah Tech University while still attending high school. High school is a subset of Non-degree-seeking. If the sql output is TRUE then the student is a high school student.

## Non-Matriculated
A Non-Matriculated student is a non-degree-seeking student who is enrolled to persue personal interests, not a credential. A Non-matriculated is a subset of Non-degree-seeking. If the sql output is TRUE then the student is a non-matriculated student.

## Online Program Student
An Online Program Student is a student who is paying online tuition because they are in an online only program (the program can only be completed online). If the sql output is TRUE then the student is an online program student.

## Overall Cumulative GPA Band
Overall Cumulative GPA Band displays a student's Overall (both institutional and transfer) Cumulative (across all semesters) GPA. Reported as a grouping: 

- new (null - no GPA yet)
- 0-1.999
- 2-2.49
- 2.5-2.999
- 3-4.0

## Institutional Cumulative GPA Band
Institutional Cumulative GPA Band displays a student's Institutional Cumulative (across all semesters) GPA. Reported as a grouping: 

- new (null - no GPA yet)
- 0-1.999
- 2-2.49
- 2.5-2.999
- 3-4.0

## Pell Eligible
A Pell Eligible student is a student who is eligible to receive a Pell grant based on their FAFSA. If the sql output is TRUE then the student is Pell eligible.

## Pell Awarded
A Pell Awarded student is a student who has been awarded a Pell grant. If the sql output is TRUE then the student is Pell Awarded.

## Subsidized Loan Awarded
A Subsidized Loan Awarded student is a student who has been awarded a Subsidized Loan. If the sql output is TRUE then the student is Subsidized Loan Awarded.

## Student Athlete Term
A Student Athlete Term is a student who was on the official team roster on or after the first scheduled contest or date of NCAA competition that semester/term. If the sql output is TRUE then the student is an athlete.


## SQL for Term Based Groups

```sql
SELECT a.student_id,
       b.college_abbrv,
       b.college_desc,
       b.department_id,
       b.department_desc,
       a.primary_program_id,
       a.primary_degree_desc AS degree_desc,
       b.major_id,
       b.major_desc,
       a.is_degree_seeking,
       a.student_type_code,
       a.student_type_desc,
       a.level_id,
       a.level_desc,
       a.primary_level_class_id,
       a.primary_level_class_desc,
       a.full_time_part_time_code AS credit_load,
       a.residency_code, 
       a.residency_code_desc,
       a.residency_in_state_desc,
       CASE
           WHEN a.student_type_code = 'H' THEN TRUE
           ELSE FALSE
        END AS is_high_school,
       CASE
           WHEN a.student_type_code = 'P' THEN TRUE
           ELSE FALSE
        END AS is_non_matriculated,
       a.is_online_program_student,
       CASE
           WHEN a.overall_cumulative_gpa < 2.0 THEN '0_to_1.999'
           WHEN a.overall_cumulative_gpa BETWEEN 2.0 AND 2.499 THEN '2_to_2.499'
           WHEN a.overall_cumulative_gpa BETWEEN 2.5 AND 2.999 THEN '2.5_to_2.999'
           WHEN a.overall_cumulative_gpa BETWEEN 3.0 AND 4 THEN '3_to_4'
           ELSE 'new'
        END AS overall_cumulative_gpa_band,
        CASE
           WHEN a.overall_institutional_gpa < 2.0 THEN '0_to_1.999'
           WHEN a.overall_institutional_gpa BETWEEN 2.0 AND 2.499 THEN '2_to_2.499'
           WHEN a.overall_institutional_gpa BETWEEN 2.5 AND 2.999 THEN '2.5_to_2.999'
           WHEN a.overall_institutional_gpa BETWEEN 3.0 AND 4 THEN '3_to_4'
           ELSE 'new'
        END AS gpa_band,
       a.is_pell_eligible,
       a.is_pell_awarded,
       a.is_subsidized_loan_awarded,
       a.is_athlete AS student_athlete_term
  FROM export.student_term_level a
  LEFT JOIN export.academic_programs b ON b.program_id = a.primary_program_id
 WHERE a.is_enrolled IS TRUE
   AND a.is_primary_level IS TRUE
   AND a.is_current_term IS TRUE;
```


# Course Based Groups
Data in this classification can change each semester and is based on courses the student is enrolled in. This data is stored at the student term section level, in the student_term_section table. This data is changes often, and can be pulled at Current, End of Term or Census.

## High School Concurrent Enrollment Student
A High School Concurrent Enrollment student is a high school student enrolled in at least one funded concurrent enrollment class. High School Concurrent Enrollment students are a subset of High School Students. If a high school student isn't a concurrent enrollment student then the student is a High School Student, Other. If the sql output is TRUE then the student is a high school concurrent enrollment student.

## SQL for High School Concurrent Enrollment Student
```sql
WITH cte AS
           (
               SELECT a.budget_code,
                      b.student_type_code,
                      a.student_id,
                      a.course_id,
                      a.term_id
                 FROM export.student_section a
                          LEFT JOIN export.student_term_level b
                                    ON b.student_id = a.student_id
                                        AND b.term_id = a.term_id
                                        AND b.is_primary_level IS TRUE
                WHERE a.is_enrolled IS TRUE
                  AND a.is_current_term IS TRUE
                ORDER BY a.student_id
           ),
       concurrent_cte AS ( --this query flags student courses that are eligible for concurrent enrollment funds
           SELECT a.student_id,
                  a.term_id,
                  CASE
                      WHEN a.student_type_code = 'H' AND (a.budget_code IN ('BC', 'SF'))
                          THEN TRUE
                      ELSE FALSE
                      END AS cte_concurrent_enrollment
             FROM cte a
       )
SELECT a.student_id,
       a.term_id,
       bool_or(cte_concurrent_enrollment) AS is_concurrent_enrollment
  FROM concurrent_cte a
 GROUP BY a.student_id, a.term_id
```

## Distance Education Status
Distance Education Status determines if a students courses are Distance Education All, Distance Education Some, or Distance Education None. A distance education course uses one or more types of technology to deliver instruction to students who are separated from the instructor and to support regular and substantive interaction between the students and the instructor synchronously or asynchronously.

## SQL for Distance Education Status

```sql
SELECT a.student_id,
       a.is_distance_ed_all,
       a.is_distance_ed_some,
       a.is_distance_ed_none
  FROM export.student_term_level a
 WHERE a.is_enrolled IS TRUE
   AND a.is_primary_level IS TRUE
   AND a.is_current_term IS TRUE
 ```

## Online Status
Online Status determines if a students courses are Online All, Online Some, or Online None. An online course is a subset of distance education courses. Online courses are primarily delivered asynchronously; students work at their own pace within the timelines established by the instructor. 

## SQL for Online Status

```sql
SELECT student_id,
       instruction_method_code_aggregated,
       instruction_method_code_aggregated !~ '[I]'  AS online_none,
       instruction_method_code_aggregated !~ '[^I]' AS online_all,
       CASE
           WHEN instruction_method_code_aggregated !~ '[^I]' THEN FALSE
           WHEN instruction_method_code_aggregated ~ '[I]+' THEN TRUE
           ELSE FALSE
           END                                      AS online_some,
       academic_year_code
FROM (SELECT a.student_id,
             STRING_AGG(instruction_method_code :: VARCHAR, ''
                        ORDER BY instruction_method_code) AS instruction_method_code_aggregated,
             b.academic_year_code
      FROM export.student_section a
               LEFT JOIN export.term b ON b.term_id = a.term_id
      WHERE is_enrolled
        AND season != 'Summer'
      GROUP BY a.student_id,
               b.academic_year_code) x
   ```

## Student Category
A student's category is based off of their degree seeking status. The options are:

- Degree Seeking
- Non-Matriculated
- High School Concurrent Enrollment Students
- High School Students, Other

## SQL for Student Category
```sql
WITH cte_degree_seeking AS
         (SELECT a.student_id,
                 a.term_id,
                 a.is_degree_seeking
          FROM export.student_term_level a
          WHERE a.is_enrolled IS TRUE
            AND a.is_primary_level IS TRUE
            AND a.is_current_term IS TRUE
          ORDER BY a.student_id),
     cte_non_matriculated AS
         (SELECT a.student_id,
                 a.term_id,
                 CASE
                     WHEN a.student_type_code = 'P' THEN 'TRUE'
                     ELSE 'FALSE'
                     END AS is_non_matriculated
          FROM export.student_term_level a
          WHERE a.is_enrolled IS TRUE
            AND a.is_primary_level IS TRUE
            AND a.is_current_term IS TRUE
          ORDER BY a.student_id),
     cte_high_school AS
         (SELECT a.student_id,
                 a.term_id,
                 CASE
                     WHEN a.student_type_code = 'H' THEN 'TRUE'
                     ELSE 'FALSE'
                     END AS is_high_school
          FROM export.student_term_level a
          WHERE a.is_enrolled IS TRUE
            AND a.is_primary_level IS TRUE
            AND a.is_current_term IS TRUE
          ORDER BY a.student_id),
     cte_concurrent AS
         (WITH cte AS
                   (SELECT a.budget_code,
                           b.student_type_code,
                           a.student_id,
                           a.course_id,
                           a.term_id
                    FROM export.student_section a
                             LEFT JOIN export.student_term_level b
                                       ON b.student_id = a.student_id
                                           AND b.term_id = a.term_id
                                           AND b.is_primary_level IS TRUE
                    WHERE a.is_enrolled IS TRUE
                      AND a.is_current_term IS TRUE
                    ORDER BY a.student_id),
               concurrent_cte AS ( --this query flags student courses that are eligible for concurrent enrollment funds
                   SELECT a.student_id,
                          a.term_id,
                          CASE
                              WHEN a.student_type_code = 'H' AND (a.budget_code IN ('BC', 'SF'))
                                  THEN TRUE
                              ELSE FALSE
                              END AS cte_concurrent_enrollment
                   FROM cte a)
          SELECT a.student_id,
                 a.term_id,
                 bool_or(cte_concurrent_enrollment) AS is_concurrent_enrollment
          FROM concurrent_cte a
          GROUP BY a.student_id, a.term_id)
SELECT a.student_id,
       a.term_id,
       b.is_degree_seeking,
       c.is_non_matriculated,
       d.is_high_school,
       e.is_concurrent_enrollment,
       CASE WHEN b.is_degree_seeking = 'TRUE' THEN 'degree-seeking'
            WHEN c.is_non_matriculated = 'TRUE' THEN 'non-matriculated'
            WHEN d.is_high_school = 'TRUE' AND e.is_concurrent_enrollment = 'TRUE' THEN 'high_school_concurrent'
            WHEN d.is_high_school = 'TRUE' AND e.is_concurrent_enrollment = 'FALSE' THEN 'high_school_other'
            ELSE 'error'
            END AS student_categories
FROM export.student_term_level a
         LEFT JOIN cte_degree_seeking b
                   ON b.student_id = a.student_id AND b.term_id = a.term_id
         LEFT JOIN cte_non_matriculated c
                   ON c.student_id = a.student_id AND c.term_id = a.term_id
         LEFT JOIN cte_high_school d
                   ON d.student_id = a.student_id AND d.term_id = a.term_id
         LEFT JOIN cte_concurrent e
                   ON e.student_id = a.student_id AND e.term_id = a.term_id
WHERE a.is_enrolled IS TRUE
  AND a.is_primary_level IS TRUE
  AND a.is_current_term IS TRUE
```
