---
title: "groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(utDataStoR)
```

Groups define the ways students are categorized. These groups can be placed in several classifications.

# Demopgraphic Groups
Data in this classification is collected from the student at the time of their application for admission. This data is stored at the student level, in the student table.

## Gender
The biological sex on a persons birth certificate. The options are M (Male) or F (Female).


## IPEDS Race/Ethnicity
Data regarding the student's race and ethnicity. There are nine options: Hispanic; American Indian/Alaskan; Asian; Black/African American; Hawaiian/Pacific Islander; White; Multiracial; Non-Resident Alien (for those on visa's); Unspecified.

## Citizenship
The student's relationship to the United States. The options are: US Citizen; Non-Res Alien-Visa or temp bas; Res Alien-immigrant status; Non-Citizen National of US; Nonimmigrant Alien 53B-8-106; Deferred Action; Student curr defined as Other.

## International
A student who is from a foreign country (has a citizenship code of Non_Res Alien-Visa) and comes to the US to take classes on a temporary F-1 visa. This is a subset of Non-Resident Alien students. If TRUE then the student is an international student.

## Age
How old a student is based on their birth date. This is typically reported in an age band: <18; 18-24; 25-34; 35-44; 44-59; >= 60.

## First Generation
A student where neither parent nor guardian has received their bachelor's degree. If TRUE then the student is a first generation student. Students who are unknown are FALSE along with those who are FALSE.

## Veteran
A student who has served in the armed forces. This data includes anyone who is receiving veteran benefits (including family members). If TRUE then the student is a veteran.

## Athlete
A student who was EVER on the official team roster on or after the first scheduled contest or date of NCAA competition. If TRUE then the student is an athlete.

## SQL for Demographic Groups

```sql
SELECT a.student_id,
       a.gender_code,
       a.ipeds_race_ethnicity,
       a.us_citizenship_desc,
       a.is_international,
       CASE
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) < 18 THEN 'less than 18'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 18 and 24 THEN '18 to 24'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 25 and 34 THEN '25 to 34'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 35 and 44 THEN '35 to 44'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) BETWEEN 45 and 59 THEN '45 to 59'
              WHEN EXTRACT(YEAR from AGE(current_date, a.birth_date)) > 59 THEN '60 plus'
              ELSE 'error'
          END AS age_band,
       a.is_first_generation,
       a.is_athlete,
       a.is_veteran
  FROM export.student a
```


# Term Based Groups
Data in this classification can change each semester and is based on information from the system for this current term such as program, credit hours attempted, etc. This data is stored at the student term level, in the student_term_level table.

## College
An organizational structure that houses departments, faculty, courses and programs. Data based off of the program the student has declared. There are nine options: College of Business; College of Education; College of Fine Arts; College of Global and Community Outreach; College of Health Sciences; College of Humanities and Social Sciences; College of Science, Engineering and Technology; and University College.

## Department
An organizational structure that is a sub-unit of a college and houses faculty, courses and programs. Data based off of the program the student has declared.

## Degree Seeking
A students who is eligible to receive financial aid and earn a degree. Data based off of the program the student has declared. The options are yes or no. Students who are not degree seeking are non-degree seeking.

## Student Type
A student's enrollment status or registration status for the term.

## Student Level
The level of degree a student is pursing. The options are: Undergraduate or Graduate.

## Class Level
A student's hierarchical status which is classified on the basis of earned credit hours for undergraduate students. The options are: Freshman <30 credit hours; Sophomore 30-59 credit hours; Junior 60-89 credit hours; Senior 90 or more credit hours. All graduate students will be Graduate Masters.

## Full-time Part-time
The status of a students time comittment based off of how many credit hours attempted and the students level for a given semester. The options are: Full-time for Undergraduate students 12 or more credit hours, for Graduate students 9 or more credit hours; Part-time for Undergraduate students less than 12 credit hours, for Graduate students less than 9 credit hours.

## Residency
A students tuition payment structure. The options are: Resident all those paying resident tuition; Non-Resident all those paying non-resident tuition; Good Neighbor paying good neighbor tuition.

## High School
A student is who is taking a course while still attending high school.

## High School Concurrent Enrollment
A High School student enrolled in at least one approved and funded concurrent enrollment class. 

## GPA
A student's Overall Cumulative GPA. This is typically reported in a GPA band:

## Pell Elligible
A student who is eligible to receive a Pell grant based on their FAFSA.

## Athlete
A student who was on the official team roster on or after the first scheduled contest or date of NCAA competition that semester. If TRUE then the student is an athlete.

## SQL for Term Based Groups

```sql
SELECT a.student_id,
       b.college_desc,
       b.department_desc,
       a.is_degree_seeking,
       a.student_type_desc,
       a.level_desc,
       a.primary_level_class_desc,
       a.full_time_part_time_code,
       a.residency_code_desc,
       CASE
           WHEN a.student_type_code = 'H' THEN TRUE
           ELSE FALSE
        END AS is_high_school,
       a.overall_cumulative_gpa,
       a.is_pell_eligible,
       a.is_athlete
  FROM export.student_term_level a
  LEFT JOIN export.academic_programs b ON b.program_id = a.primary_program_id
 WHERE a.is_enrolled IS TRUE
   AND a.is_primary_level IS TRUE
   AND a.is_current_term IS TRUE
```

# Location Groups
Data in this grouping is collected from the students FIRST application for admission.

## Country of Origin
The country the student lived at the time fo first application.

## State of Origin
The state the student lived at the time fo first application. Students who were not in the US get grouped into Outside the 50 states/territories and the District of Columbia.

## County of Origin
The county the student lived at the time of first application. Students who were not from Utah get grouped into Out-of-State, in US; or Out-of-State, out of US.

# Course Based Groups
Data in this grouping can change each semester and is based on courses the student is enrolled in. This data is stored at the student term section level, in the student_term_section table.
