---
title: "groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(utDataStoR)
```

Groups define all the ways you can categorize students. These groups can be placed in several classifications.

# Demopgraphic Groups
Data in this grouping is collected from the student at the time of their application for admission.

## Gender
The biological sex on a persons birth certificate. The options are Male or Female.

## IPEDS Race/Ethnicity
Data regarding the student's race and ethnicity. There are nine options: Hispanic; American Indian or Alaska Native; Asian; Black or African American; Native Hawaiian or Other Pacific Islander; White; Multiracial; Nonresident (for those on visa's); Unspecified. 

## Citizenship
The student's relationship to the United States. The options are: US Citizen; Non-Resident Alien; Resident Alien or Other Eligible Non-citizens; Non-immigrant Undocumented Students; US Non-citizen Nationals; Other.

## International
A student who is from a foreign country and comes to the US to take classes on a temporary F-1 visa. This is a subset of Non-Resident Alien students.

## Age
How old a student is based on their birth date. This is typically reported in an age band: <18; 18-24; 25-34; 35-44; 44-59; >= 60.

## First Generation
A student where neither parent nor guardian has received their bachelor's degree. The options are yes, no. Those who do not answer the question on the application are listed as no.

## Veteran
A student who has served in the armed forces. This data may include anyone who is receiving veteran benefits (including family members).

## Athlete
A student on the official team roster on or after the first scheduled contest or date of NCAA competition.

## High School
A student is who is taking a course while still attending high school.


# Term Based Groups
Data in this grouping can change each semester and is based on information from the system such as program, credit hours attempted, etc.

## College
An organizational structure that houses departments, faculty, courses and programs. Data based off of the program the student has declared. There are nine options: College of Business; College of Education; College of Fine Arts; College of Global and Community Outreach; College of Health Sciences; College of Humanities and Social Sciences; College of Science, Engineering and Technology; and University College.

## Department
An organizational structure that is a sub-unit of a college and houses faculty, courses and programs. Data based off of the program the student has declared.

## Degree Seeking
A students who is eligible to receive financial aid and earn a degree. Data based off of the program the student has declared. The options are yes or no. Students who are not degree seeking are non-degree seeking.

## Student Type
A student's enrollment status or registration status for the term.

## Student Level
The level of degree a student is pursing. The options are: Undergraduate or Graduate.

## Class Level
A student's hierarchical status which is classified on the basis of earned credit hours for undergraduate students. The options are: Freshman <30 credit hours; Sophomore 30-59 credit hours; Junior 60-89 credit hours; Senior 90 or more credit hours. All graduate students will be Graduate Masters.

## Full-time Part-time
The status of a students time comittment based off of how many credit hours attempted and the students level for a given semester. The options are: Full-time for Undergraduate students 12 or more credit hours, for Graduate students 9 or more credit hours; Part-time for Undergraduate students less than 12 credit hours, for Graduate students less than 9 credit hours.

## Residency
A students tuition payment structure. The options are: Resident all those paying resident tuition; Non-Resident all those paying non-resident tuition; Good Neighbor paying good neighbor tuition. 

## High School Concurrent Enrollment
A High School student enrolled in at least one approved and funded concurrent enrollment class. 

## GPA
A student's Overall Cumulative GPA. This is typically reported in a GPA band:

## Pell Elligible
A student who is eligible to receive a Pell grant based on their FAFSA.


# Location Groups
Data in this grouping is collected from the students FIRST application for admission.

## Country of Origin
The country the student lived at the time fo first application.

## State of Origin
The state the student lived at the time fo first application. Students who were not in the US get grouped into Outside the 50 states/territories and the District of Columbia.

## County of Origin
The county the student lived at the time of first application. Students who were not from Utah get grouped into Out-of-State, in US; or Out-of-State, out of US.
