---
title: "groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utDataStoR)
```

Groups defines all the ways you can group or categorize students. These groupings can be placed in several categories are demographic information but some will be about the types of courses the students are enrolled in.

# Demopgraphic Groups
Data in this grouping is collected from the student at the time of their application for admission.

## Gender
Data regarding the student's gender. The options are Male or Female.

## IPEDS Race/Ethnicity
Data regarding the student's race and ethnicity. There are nine options: Hispanic; American Indian or Alaska Native; Asian; Black or African American; Native Hawaiian or Other Pacific Islander; White; Multiracial; Nonresident (for those on visa's); Unspecified. 

## Age
Data based off of the student's birth date. This is reported in an age band typically: <18; 18-24; 25-34; 35-44; 44-59; >= 60.

## First Generation
Students where neither parent completed a Bachelor's degree. The options are yes, no. Those who do not answer the question on the application are listed as no.

## Veteran
?? Work on this.

# Term Based Groups
These are groups that can change each semester. They are based on data from the system such as program, credit hours attempted, etc.

## College
An organizational structure that houses departments, faculty, courses and programs. Data based off of the program the student has declared. There are nine options: College of Business; College of Education; College of Fine Arts; College of Global and Community Outreach; College of Health Sciences; College of Humanities and Social Sciences; College of Science, Engineering and Technology; and University College.

## Department
An organizational structure that is a sub-unit of a college and houses faculty, courses and programs. Data based off of the program the student has declared.

## Degree Seeking
Students who are eligible to receive financial aid and earn a degree. Data based off of the program the student has declared. The options are yes or no. Students who are not degree seeking are non-degree seeking.

## Student Type
A student's enrollment status or registration status for the term.

## Student Level


## Class Level
Based off of how many credit hours a student has completed. The options are: Freshman <30 credit hours; Sophomore 30-59 credit hours; Junior 60-89 credit hours; Senior 90 or more credit hours; Graduate graduate masters.
