---
title: "Headcount"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{headcount}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utDataStoR)
```

Headcount is the measure of the net unduplicated number of students enrolled at Utah Tech for a specific period of time. Headcount is used for reporting and inquiry.

## Point In Time Headcount

The point in time headcount is a snapshot of current net unduplicated enrollment data for each calendar day during an academic term. 

Point in time headcount data returns the academic term, date, week of the year, the number of days to official term start date, and the total headcount snapshot.

```sql
SELECT term_id,
       date,
       date_part('week', date) - 11 AS week,
       days_to_class_start,
       COUNT(DISTINCT student_id) AS headcount
FROM export.daily_enrollment
WHERE term_id = '202340'
AND is_enrolled IS TRUE
GROUP BY term_id, date, days_to_class_start
ORDER BY date DESC;
```

## Current Headcount

The current net unduplicated headcount is the number of students enrolled for an academic term.

Current headcount can change daily for an academic term as it is active data.

The current headcount may also change after the semester ends if official changes are made to student record such as withdrawal status.

```sql
   SELECT a.term_desc,
          COUNT(a.sis_system_id) AS current_headcount
     FROM export.student_term_level_version a
LEFT JOIN export.term b
       ON a.term_id = b.term_id
    WHERE a.is_enrolled IS TRUE
      AND a.is_primary_level IS TRUE
      AND a.version_desc = 'Current'
      AND DATE_PART('year', NOW()) - b.academic_year_code :: INT <= 5 -- Current term plus last 5 Years.
 GROUP BY a.term_desc
 ORDER BY a.term_desc;
``` 

## Census Headcount

The census headcount is the net unduplicated number of students enrolled as of the 15th day of instruction of an academic term. Census headcount is based on snapshot data, this headcount returns a static headcount and is used for reporting.

```sql
...
AND a.version_desc = 'Census' --Version description of snapshot status.
...
```

## End of Term Headcount

The end of term headcount is the net unduplicated number of students enrolled at the end of the semester and after the high school concurrent enrollment grades are entered. Based on snapshot data, this headcount returns a static headcount and is used for reporting.

```sql
...
AND a.version_desc = 'End of Term' --Version description of snapshot status.
...
```

## IPEDS Fall Enrollment Headcount

The Fall Enrollment survey is based on a snapshot of Utah Tech's unduplicated Fall census enrollment headcount and is used for reporting annual IPEDS Fall Enrollment.

```sql
...
AND a.version_desc = 'Census' --Version description of snapshot status.
...
```

## IPEDS 12 Month Enrollment Headcount

The purpose of the 12-month Enrollment component of IPEDS is to collect unduplicated census enrollment counts of all students enrolled for credit and instructional activity data a 12-month period. 

```sql
  SELECT b.academic_year_code,
         COUNT(DISTINCT a.sis_system_id)
     FROM export.student_term_level_version a
LEFT JOIN export.term b
       ON a.term_id = b.term_id
    WHERE a.is_enrolled = TRUE
      AND a.is_primary_level = TRUE
      AND a.version_desc = 'Census'
      AND b.academic_year_code = '2022'
      AND b.season != 'Summer'
 GROUP BY academic_year_code;
 ```

## USHE Headcount

The USHE headcount reports unduplicated net enrollment headcount five times per fiscal year:

- Summer End of Term
- Fall Census 
- Fall End of Term
- Spring Census
- Spring End of Term

**Census**: USHE Census Headcount is the net unduplicated number of students enrolled as of the 15th day of instruction of an academic term. Census headcount is based on snapshot data, this headcount returns a static headcount and may be modified to meet USHE reporting element criteria.

```sql
...
AND a.version_desc = 'Census' --Version description of snapshot status.
...
```

**End of Term**: USHE End of Term Headcount is the net unduplicated number of students enrolled at the end of the semester and after the high school concurrent enrollment grades are entered. Based on snapshot data, this headcount returns a static headcount and may be modified to meet USHE reporting element criteria.

```sql
...
AND a.version_desc = 'End of Term' --Version description of snapshot status.
...
```




