---
title: "Headcount"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{headcount}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utDataStoR)
```

Headcount is a measure of the net unduplicated number of students enrolled at Utah Tech for a specific period of time and how it is used for official reporting and inquiry.

This vignette will focus on the metrics involved in measuring net headcount and how they differ.

## Current Headcount

The current net unduplicated headcount is the number of students enrolled. 

- Example: The current headcount for Fall 2021 is 12,266 students who are enrolled at UT. This headcount can vary by day during an academic term as it is active data.

## Census Headcount

The census headcount is the net unduplicated number of students who were enrolled as of the 15th day of instruction of an academic term. This headcount is set as a static number that will not change as it is based on snapshot data and used as an official reporting metric.

- Example: The census headcount for Fall 2021 is 12,266.

## End of Term Headcount

The end of term headcount is the net unduplicated number of students who were enrolled as of the final day of instruction of an academic term. This headcount is set as a static number that will not change as it is based on snapshot data and used as an official reporting metric.

- Example: The end of term headcount for Fall 2021 is 12,277.

## IPEDS Headcount

The Fall Enrollment survey is based on a snapshot of Utah Tech's unduplicated fall census enrollment headcount. This headcount aligns with Fall census headcount.

- Example: IPEDS Fall Enrollment 2021-22 is 12,266. 


## USHE Headcount

The USHE Students Data Submission File reports net headcount by extract five times per fiscal year:

- Summer End of Term
- Fall Census 
- Fall End of Term
- Spring Census
- Spring End of Term

**Census**: USHE Students Data Submission File is based on a snapshot of Utah tech's unduplicated academic term enrollment headcount. 

The Fall Census Students Data Submission file aligns with the IPEDS Fall Enrollment Survey. 

- Example: The USHE Fall Census Students Data Submission File headcount for Fall 2021 is 12,266

Reporting for the Spring Census Students Data Submission file is based on the static net enrollment headcount for the 3rd week or 15th day of instruction for a Spring academic term.

**End of Term**: USHE Students Data Submission File

The end of term headcount is the net unduplicated number of students who were enrolled as of the final day of instruction of an academic term. This headcount is set as a static number that will not change as it is based on snapshot data and used as an official reporting metric.

- Example: The end of term headcount for Fall 2021 is 12,277.


