---
title: "utDataStoR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{utDataStoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package is designed to provide a collection of simple, standardized and approved R code snippets and functions that are commonly used across various projects at Utah Tech University. utDataStoR is greatly enhanced when used in tandem with [utHelpR](https://github.com/ut-effectiveness/utHelpR) package.


Follow the steps below to get started:

### Loading the Package

Once installed, load the package into your R session:

```{r, eval=FALSE}
library(utHelpR)
library(utDataStoR)
```

### Using the Functions

The package includes a variety of functions that you can use in your projects. Here are a few examples:

### Example Function 1: make_headcount_sql

First you will want to make sure you are choosing the correct sql you can reference
documentation in the **Articles** tab. In this case you would want to look at `vignette("headcount")`

Your project must have a folder called **sql** you can make one by hand or use the utHelpR package 
```{r, eval=FALSE}
utHelpR::make_standard_folders()
```

```{r, eval=FALSE}
# Example usage of a make_headcount_sql function
utDataStoR::make_headcount_sql("headcount.sql", "census")
```

This will write the desired sql script (**headcount.sql**) to a .sql file housed in sql/ folder

you could then run the sql by using the utHelpR package

```{r, eval=FALSE}
head_count_df <- utHelpR::get_data_from_sql_file("headcount.sql")
```

### Example Function 2: display_sql_with_comments

```sql
`r utDataStoR::display_sql_with_comments("headcount", "headcount_census.sql")`
```

### Documentation

Detailed documentation for each function is available on the pkgdown site. You can also access the documentation within R using the `?` help operator:
`?make_headcount_sql` or `?display_sql_with_comments`


### Contributing

We welcome contributions from the Utah Tech University community! If you have suggestions for new functions or improvements to existing ones, participate in [GitHub/discussion](https://github.com/ut-effectiveness/utDataStoR/discussions) please submit a pull request on our GitHub repository.

### Getting Help

If you encounter any issues or have questions, please open an issue on [GitHub/issues](https://github.com/ut-effectiveness/utDataStoR/issues) or contact the package maintainer at [datablaze\@utahtech.edu](mailto:datablaze@utahtech.edu){.email}.

------------------------------------------------------------------------

Feel free to customize this guide further to better fit your specific needs! If you have any other questions or need additional help, just let me know.
