---
title: "Fake Retention Data"
output: html_document
date: "2022-07-27"
---

```{r}
library(here)
library(stringr)
library(tidyverse)
library(rlang)
library(utHelpR)
```

```{r}
ipeds <- c(
  "White",
  "Hispanic",
  "Multiracial",
  "Black/African American",
  "Unspecified",
  "American Indian/Alaskan",
  "Non-Resident Alien",
  "Asian",
  "Hawaiian/Pacific Islander"
  )

gpa_bands <- c(
  'below 2.0',
  '2.0 to 2.9',
  '3.0 to 4.0'
)

retention_status <- c(
  'spring returned',
  'fall returned',
  'third fall returned',
  'fourth fall returned',
  'fifth fall returned',
  'sixth fall returned',
  'stopped out',
  'dropped out', 
  'graduated'
)
```

```{r}
program_college_pull <- get_data_from_sql_file(
  here::here('sandbox', 'fake_data', 'sql', 'program_college.sql'),
  'edify',
  'none') 

program_college_missing_values <- program_college_pull 

program_college_missing_values[16, 3] <- 'UCOL'
program_college_missing_values[16, 4] <- 'University College'
program_college_missing_values[26, 3] <- 'GCO'
program_college_missing_values[26, 4] <- 'Global & Community Outreach'

program_college <- program_college_missing_values %>% 
  filter(!is.na(department_id)) %>% 
  arrange(college_abbrv)

college_list <- program_college$college_abbrv %>% unique()

department_list <- program_college$department_id %>% unique()
```


```{r}
fake_id <- function(id_length, col_length, unique) {
  # id_length is the length of the id number
  # col_length is how many id numbers you will need
  # unique is if you want each id to be unique


  # use in integer_list to generate list of integers to sample from
  finish_num <- paste(rep(9, id_length), collapse = "")
  # random selection of integers
  integer_list <- sample(1:finish_num, col_length, replace = unique)
  # format the integers
  stringr::str_pad(integer_list, id_length, pad = '0')
}

# Create true false columns
tf <- function(sample_size) {
  sample(c(TRUE, FALSE), sample_size, replace = TRUE)
}

# put the columns in alphabetic order
order_cols <- function(data_df) {
  data_df %>%
    select(order(colnames(data_df)))
}
```

## Third and fourth tab

The data frame for the third and fourth tabs consists of a population characteristics data frame combined with a collection of return metric data frames. Each metric data frame is combined with the population characteristics data frame by combining the columns. We form the total metric by combining all the rows of each of the resulting data frames. 


```{r}
cohort_years <- as.character(2017:2022)
cohort_population <- c(8145, 8934, 9845, 10400, 12345, 13082)
total_cohort_population <- sum(cohort_population)



population_characteristics_df <- tibble(
  student_id = fake_id(8, total_cohort_population, unique = TRUE),
  cohort = c(
    rep('2017', cohort_population[[1]]),
    rep('2018', cohort_population[[2]]),
    rep('2019', cohort_population[[3]]),
    rep('2020', cohort_population[[4]]),
    rep('2021', cohort_population[[5]]),
    rep('2022', cohort_population[[6]])
    ),
  college = sample(college_list, total_cohort_population, replace = TRUE),
  department = sample(department_list, total_cohort_population, replace = TRUE),
  program = department,
  ipeds_race_ethnicity = sample(ipeds, total_cohort_population, replace = TRUE),
  gender = tf(total_cohort_population)
)

make_return_metrics <- function(metric) {
  
  metric_df <- tibble(
    spring_returned = tf(total_cohort_population),
    fall_returned = tf(total_cohort_population),
    second_fall_returned = tf(total_cohort_population),
    third_fall_returned = tf(total_cohort_population),
    fourth_fall_returned = tf(total_cohort_population),
    fifth_fall_returned = tf(total_cohort_population),
    sixth_fall_returned = tf(total_cohort_population),
    gpa_band = sample(gpa_bands, total_cohort_population, replace = TRUE),
    metric = metric
    )
  
  bind_cols(
    population_characteristics_df,
    metric_df
  )
}

tabs_3_and_4_df <- bind_rows(
  make_return_metrics('third week'),
  make_return_metrics('USHE'),
  make_return_metrics('end of term'),
  make_return_metrics('ipdes')
)
```

## First tab

```{r}
make_historical_rates <- function(start, slope) {
  
  start = 50
  
  slope = 6
  
  x <- 0:5
  
  noise <- rnorm(6, 0, 10)
  
  line <- start + x*slope  
  
  return_rate <- line + noise %>% 
    round(digits = 1)
  
  return_rate
}

make_hist_rate_tibble <- function(metric, start, slope) {
  tibble(
    year = cohort_years,
    metric = rep(metric, 6),
    return_rate = make_historical_rates(start, slope)
  )
}

tab_1_df <- bind_rows(
  make_hist_rate_tibble('fall to fall', 50, 6),
  make_hist_rate_tibble('fall to spring', 40, 3),
  make_hist_rate_tibble('second year', 60, 3),
  make_hist_rate_tibble('third year', 50, -5),
)

tab_1_goals <- tibble(
  metric = c('fall to fall', 'fall to spring', 'second year', 'third year'),
  goal = c(50, 60, 74, 89)
) 

```

## Save
```{r}
save(
  tab_1_df,
  tab_1_goals,
  tabs_3_and_4_df,
  file = here('sandbox', 'fake_data', 'fake_data.rda')
)
```






